Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-01-10 ms/manuscript.Rmd
Use unicode lambda instead of latex

We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
$\lambda$ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
\K describes the ratio between the amount of observed variance in traits vs. the amount of variance expected under BM: when $\K > 1$, traits are more conserved than expected under BM; when $\K < 1$, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in $\lambda$ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of $\lambda$ vs. $\lambda = 0$ (no phylogenetic signal).
We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010].
Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits.
We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ("strict" dataset).

@@

Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for $\lambda$ and K.
When measured with $\lambda$, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of $\lambda$ close to 1; only number of pinnae pairs had $\lambda$ close to zero (\tab \@ref(tab:phylosig-cont)).
However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of $\K < 1$ (\tab \@ref(tab:phylosig-cont)).
All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
Similar results were obtained using the "strict" dataset (\tab S4).

@@

# Include in PDF in latex format
phylosig_cont_traits_table %>%
  rename(K = kval, `P(K)` = k_pval, `P(lambda)` = lambda_pval) %>%
  column_to_rownames("trait") %>%
  kable(caption = "Phylogenetic signal in continuous traits", format = "latex", booktabs = TRUE) %>%
  footnote(general = "K, Blomberg's \\\\textit{K} (Blomberg \\\\textit{et al.}, 2003); lambda, Pagel's lambda (Pagel, 1999). Values of lambda or \\\\textit{K} greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or K near 0 indicate random distribution of traits.", general_title = "", threeparttable = TRUE, escape = FALSE)
```

>>>>

We tested for phylogenetic signal in quantitative traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999], two metrics with slightly different interpretations and sensitivity to different evolutionary scenarios [@Munkemuller2012], with the \function{phylosig} function in the \package{phytools} R package [@Revell2012].
λ is a scaling parameter that ranges from 0 (traits evolving randomly) to 1 (traits evolving along branches according to Brownian motion (BM)).
\K describes the ratio between the amount of observed variance in traits vs. the amount of variance expected under BM: when \K > 1, traits are more conserved than expected under BM; when \K < 1, traits have less phylogenetic signal than expected under BM [@Blomberg2003].
We tested for significance by comparing observed \K against a null distribution of \num{1000} trees with trait values randomized across the tips, and in λ by conducting a likelihood ratio test comparing the log likelihoods of the observed value of λ vs. λ = 0 (no phylogenetic signal).
We tested for phylogenetic signal in qualitative traits using Fritz and Purvis' *D* [@Fritz2010].
Values of \D range from 0 under a model of trait evolution by BM, to 1 for random distribution of traits.
We analyzed \D using the \function{phylo.d} function in the \package{caper} R package [@Orme2013], and conducted significance test by comparing observed \D to distributions of simulated values produced under two models: random shuffling of traits on the tree, or simulation of a binary trait under a BM model.
To account for possible circularity in measuring the phylogenetic signal of traits that were scored based on taxonomy, we conducted the same analyses on a dataset excluding species that had been scored this way ("strict" dataset).

@@

Most measured traits showed some degree of phylogenetic signal, but the strength varied across traits (\tabs \@ref(tab:phylosig-cont), \@ref(tab:phylosig-binary), \fig \@ref(fig:tree)), and for quantitative traits, different results were obtained for λ and K.
When measured with λ, most sporophyte traits show phylogenetic signal (as expected under a BM model), with values of λ close to 1; only number of pinnae pairs had λ close to zero (\tab \@ref(tab:phylosig-cont)).
However, when measured with \K, only frond width and rhizome diameter showed strong phylogenetic signal (more than expected under BM); other traits had values of \K < 1 (\tab \@ref(tab:phylosig-cont)).
All binary gametophyte traits showed significant phylogenetic signal, either similar to (hairs, \D = `r hairs_d`) or more conserved (gemmae, \D = `r gemmae_d`; glands, \D = `r glands_d`; morphotype, \D = `r morphotype_d`) than expected under BM (\tab \@ref(tab:phylosig-binary)).
Growth habit showed phylogenetic signal nearly identical to that expected under BM (\D = `r habit_d`; \tab \@ref(tab:phylosig-binary)).
Similar results were obtained using the "strict" dataset (\tab S4).

@@

# Include in PDF in latex format
phylosig_cont_traits_table %>%
  rename(K = kval, `P(K)` = k_pval, `P(lambda)` = lambda_pval) %>%
  column_to_rownames("trait") %>%
  kable(caption = "Phylogenetic signal in continuous traits", format = "latex", booktabs = TRUE) %>%
  footnote(general = "K, Blomberg's \\\\textit{K} (Blomberg \\\\textit{et al.}, 2003); lambda, Pagel's lambda (Pagel, 1999). Values of lambda or \\\\textit{K} greater than 1 indicate traits evolving according to Brownian Motion (BM); values of lambda or \\\\textit{K} near 0 indicate random distribution of traits.", general_title = "", threeparttable = TRUE, escape = FALSE)
```
